The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-10-09 Paper.qmd

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child variables.

@@

::: {#fig-PCA}
```{r}
plot_pca
```

A) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable. B) Loadings of PC1 and PC2 on each variable.
:::

@@

Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week. However, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents. See @fig-barplot_ConflictFreq.

The frequency of alloparenting effort was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day. See @fig-barplot_AlloparentingFreq.

>>>>

<!-- [NOTE: Add categorical variables to the table; combine child and adult variables into 1 table] -->
```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child variables.

@@

A PCA of primary study variables found that sociodemographic and ecological variables loaded primarily on PC1, whereas the child signaling, conflict, and child age variables loaded primarily on PC2. Child signaling variables modestly load on PC1 as well in a direction suggesting that greater signaling covaries with more adverse environments. See @fig-PCA.

::: {#fig-PCA}
```{r}
#| fig-width: 11
plot_pca
```

A) Loadings of PC1 and PC2 on each variable. B) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable. 
:::

@@

Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week. However, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents. See @fig-barplot_ConflictFreq. Conflict was also associated with higher signaling frequencies. See @fig-PCA and @fig-signal_corrplot.

The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day. See @fig-barplot_AlloparentingFreq.
